Formal Verification of Real-Time Data Processing of the LHC Beam Loss Monitoring System: A Case Study

### 2011-05-26 fmics/paper.tex
Add Naghmeh's responses to minor comments from Dropbox

Compared to test-based methods, like simulation, formal methods not only offer much higher confidence in the correctness of a system's behavior, but also help improve our understanding of its specification.
One of the challenges in pursuing a formal verification approach for SRS was capturing the intricate details of the system's specification via experiment and refinement with a team of different backgrounds and expertise.
Our confidence in the SRS design as a result of this effort ultimately rests upon our deep understanding of why the design is correct rather than the fact that we obtained ``Theorem Proved" as the final output of a software tool.
In particular, our use of mechanized logical deduction was a highly iterative process that incrementally refined our understanding of (1) the implementation (2) the intended behavior and (3) the ``whiteboard-level" argument or explanation for why the implementation achieves the intended behaviors.
The most important use of HOL4  was its role as an ``implacable skeptic" that insisted we provide justification and compelled us to clarify the details~\cite{rushby}.

@@

\emph{(a) Receive, Check, and Compare (RCC)}: The RCC block receives data directly from the acquisition cards, and attempts to detect and correct erroneous transmissions by using Cyclic Redundancy Check and 8B/10B algorithms~\cite{CRC,8B10B}.

@@

\emph{(c) Threshold Comparator}: Every running sum needs to be compared to the threshold determined by the beam energy reading at that moment.
The comparator initiates a beam dump request if any of the running sums is higher than its corresponding threshold.
Beam dump requests are forwarded to the Beam Interlock System which initiates the beam dump.
There are 12 running sums calculated for each 16-detector channel allocated to a BLETC card.
The beam energy information is scaled to 32 levels (0.45 to 7 TeV) and each processing module holds data only for those 16 connected detectors.
Thus, a total of 6,144 threshold values need to be held on each card.

@@

HOL4 is an \emph{interactive} theorem prover: the user provides the high level proof strategy by composing functions that automate common chains of logical deduction.
The work described here could have been done using other systems such as HOL Light~\cite{HOLLight}, Isabelle/HOL~\cite{Isabelle}, ProofPower~\cite{ProofPower}, PVS~\cite{PVS} or Coq~\cite{Coq}.
The first three use essentially the same higher-order logic as HOL4, whilst PVS and Coq support more powerful logics.
%<What about ACL2?>
While offering less ``push-button'' automation than other kinds of formal verification such as model-checking, machine-assisted theorem proving is appropriate for verifying the SRS, since it is unlikely that the correctness theorems, which need lemmas proved by manually guided mathematical induction, could be generated automatically.

@@

\begin{figure}[p]
\begin{tabular}{ll}
$\;\mathsf{tap}\;0\;0=0\quad\quad$ & $\mathsf{tap}\;0\;x=0$\\
$\;\mathsf{tap}\;1\;0=1-1\quad$ &  $\mathsf{tap}\;1\;x=2-1$\\
$\;\mathsf{tap}\;2\;0=8-1\quad$ &  $\mathsf{tap}\;2\;x=16-1$\\
$\;\dots\quad$ & $\mathsf{tap}\;6\;x=128-1\quad\text{(where each $x>0$)}$\\[1ex]

@@

Since the outputs of a slice stay constant between update times for the slice, Theorem~\ref{output_input_at_update_times} suffices to characterize all outputs at all times.
Thus, the SRS structure's outputs are equal to the exact running sums of the input counts over windows whose sizes depend on the the width of the slice's shift register and position of the taps.
However, the sums are delayed by the total delay across all previous slices (represented by $\mathsf{delay\_sum}$).

@@

Another future research direction is to refine our formal model to a lower level of abstraction, closer to the hardware level (e.g., by representing count values as finite words instead of natural numbers).
We are also considering verifying a model extracted from the Hardware Description Language (HDL) used to synthesize the BLETC FPGA.

>>>>

Our interest in using formal methods was originally motivated by questions about the SRS that arose in the course of an external technical audit of the BLMS performed by two of the co-authors, Ghafari and Joyce, and their colleagues.
Compared to test-based methods, like simulation, formal methods not only offer much higher confidence in the correctness of a system's behavior, but also help improve our understanding of its specification.
One of the challenges in pursuing a formal verification approach for SRS was capturing the intricate details of the system's specification via experiment and refinement with a team of different backgrounds and expertise.
Our confidence in the SRS design as a result of this effort ultimately rests upon our deep understanding of why the design is correct rather than the fact that we obtained ``Theorem Proved" as the final output of a software tool.
In particular, our use of mechanized logical deduction was a highly iterative process that incrementally refined our understanding of (1) the implementation (2) the intended behavior and (3) the ``whiteboard-level" argument or explanation for why the implementation achieves the intended behaviors.
The most important use of HOL4  was its role as an ``implacable skeptic" that insisted we provide justification and compelled us to clarify the details~\cite{rushby}.

@@

\emph{(a) Receive, Check, and Compare (RCC)}: The RCC block receives data directly from the acquisition cards, and attempts to detect erroneous transmissions by using Cyclic Redundancy Check and 8B/10B algorithms~\cite{CRC,8B10B}.

@@

\emph{(c) Threshold Comparator}: Every running sum needs to be compared to the threshold determined by the beam energy reading at that moment.
The comparator initiates a beam dump request if any of the running sums is higher than its corresponding threshold.
Beam dump requests are forwarded to the Beam Interlock System which initiates the beam dump.
There are 12 running sums calculated for each 16-detector channel allocated to a BLETC card.
There are 32 levels (0.45 to 7 TeV) of beam energy and each processing module holds data only for those 16 connected detectors.
%The beam energy information is scaled to 32 levels (0.45 to 7 TeV) and each processing module holds data only for those 16 connected detectors.
Thus, a total of 6,144 threshold values need to be held on each card.

@@

HOL4 is an \emph{interactive} theorem prover: the user provides the high level proof strategy by composing functions that automate common chains of logical deduction.
The work described here could have been done using other systems such as HOL Light~\cite{HOLLight}, Isabelle/HOL~\cite{Isabelle}, ProofPower~\cite{ProofPower}, PVS~\cite{PVS} or Coq~\cite{Coq}.
The first three use essentially the same higher-order logic as HOL4, whilst PVS and Coq support more powerful logics.
%<What about ACL2?>
While offering less ``push-button'' automation than other kinds of formal verification such as model-checking, machine-assisted theorem proving using HOL4 is appropriate for verifying the SRS, 
since it generates correctness theorems; in addition, it gives a way to very explicitly parameterize the model.
%since it is unlikely that the correctness theorems, which need lemmas proved by manually guided mathematical induction, could be generated automatically.

@@

\begin{figure}[p]
\begin{tabular}{ll}
$\;\mathsf{tap}\;0\;0=0\quad\quad$ & $\mathsf{tap}\;0\;x=0$\\
$\;\mathsf{tap}\;1\;0=1-1\quad$ &  $\mathsf{tap}\;1\;x=2-1$\\
$\;\mathsf{tap}\;2\;0=8-1\quad$ &  $\mathsf{tap}\;2\;x=16-1$\\
$\;\mathsf{tap}\;3\;0=32-1\quad$ &  $\mathsf{tap}\;3\;x=128-1$\\
$\;\mathsf{tap}\;4\;0=32-1\quad$ &  $\mathsf{tap}\;4\;x=256-1$\\
$\;\mathsf{tap}\;5\;0=16-1\quad$ &  $\mathsf{tap}\;5\;x=64-1$\\
$\;\mathsf{tap}\;6\;0=32-1\quad$ &  $\mathsf{tap}\;6\;x=128-1\quad\text{(where each $x>0$)}$\\
$\;\dots\quad$ & $\mathsf{tap}\;n\;x=0\quad\text{(where $x\geq0$ and $n>6$)}$\\[1ex]

@@

Since the outputs of a slice stay constant between update times for the slice, Theorem~\ref{output_input_at_update_times} suffices to characterize all outputs at all times.
Thus, the SRS structure's outputs are equal to the exact running sums of the input counts over windows whose sizes depend on the width of the slice's shift register and position of the taps.
However, the sums are delayed by the total delay across all previous slices (represented by $\mathsf{delay\_sum}$).

@@

Another future research direction is to refine our formal model to a lower level of abstraction, closer to the hardware level (e.g., by representing count values as finite words instead of natural numbers).
We are also considering verifying a model extracted from the Hardware Description Language (HDL) used to synthesize the BLETC FPGA.
\\

{\bf Acknowledgment:} The authors thank Mike Gordon at the University of Cambridge for his insightful and helpful feedback during the course of this work.
